Windowed least-squares model reduction for dynamical systems

### 2019-10-15 section4/spacetime_numerical.tex
minor changes to space--time solution methods

\begin{remark}
For the limiting case where the window size spans the entire domain, $\DeltaSlabArg{1}\equiv T$, and a uniform quadrature weights are used, direct \methodAcronym\ with a \spaceTimeAcronym\ trial subspace recovers ST-LSPG.
\end{remark}

>>>>

\begin{remark}
For the limiting case where the window size spans the entire domain, $\DeltaSlabArg{1}\equiv T$, and uniform quadrature weights are used, and the weighting matrix $\stweightingMatArg{n} = \mathbf{I}$, direct \methodAcronym\ with a \spaceTimeAcronym\ trial subspace recovers ST-LSPG with $\lspgWeightingSTArg{\cdot} = \mathbf{I}$. For equivalence conditions in the case where $\lspgWeightingSTArg{\cdot} \ne \mathbf{I}$, the objective function~\eqref{eq:obj_gen_lms_final_st} can be modified to enable weighted space--time norm. 
\end{remark}

### 2019-10-15 section4/spatial_numerical.tex
thru 4.1
minor changes to space--time solution methods

\subsection{\spatialAcronym: direct solution approach}\label{sec:direct} 
\KTC{Does the time grid definted in this section have any relationship to the
time grid used for time discretization? Haven't gotten past HERE below, but
wanted to check (maybe it's made clear later).}
\EP{This time grid becomes the time discretization over each window. At this point in the paper, 
there is no notion of a time grid for \methodAcronym}

Direct approaches solve~\eqref{eq:tclsrm} by
``transcribing" the infinite dimensional optimization problem into a finite
dimensional one by discretizing the state and objective functional in time.
The minimization problem is then reformulated as a (non)linear programming, or
optimization, problem. A variety of direct solution approaches exist, 
including collocation approaches, spectral  methods, and genetic algorithms.

@@

Linear multistep schemes approximate the solution at time instance $\timeWindowArg{n}{i}$ using the previous $k$ time-instances.  

@@

\methodAcronym\ sequentially computes the approximate solutions $\approxstateDiscreteArg{n,0},\ldots,\approxstateDiscreteArg{n,\nstepsArg{n}}$, $n=1,\ldots,\nslabs$, as the solution to the minimization problems

@@

\begin{remark}
For the limiting case where the window size is equivalent to the time-step size, $\DeltaSlabArg{n} \equiv \timeWindowArg{n}{1} - \timeWindowArg{n}{0}$, and uniform 
quadrature weights are used, \methodAcronym\ recovers LSPG.
\end{remark} 
\subsubsection{Solution to the least-squares problem through the Gauss--Newton method}
Discretization through linear multistep methods (as well as other techniques) 
results in a (non)linear least-squares problem.
A variety of algorithms exist for solving least-squares problems, including trust region approaches, the Gauss–Newton method, and the Levenberg–Marquardt method.  
The numerical experiments presented in this work consider the Gauss--Newton method, and as such we outline this approach here.

To describe the Gauss--Newton method, we first define a ``vectorization" function that takes a tuple of time-local solutions at each time instance and outputs 
a concatenated vector, 
\begin{align*}
 \unroll &\vcentcolon (\stateyDiscreteArg{1},\ldots,\stateyDiscreteArg{m} ) \mapsto \begin{bmatrix} [\stateyDiscreteArg{1}]^T & \ldots & [\stateyDiscreteArg{m}]^T \end{bmatrix}^T  \\
&\vcentcolon \RR{p} \otimes \RR{m} \rightarrow \RR{pm},
\end{align*}
for arbitrary $p$ and $m$. The stacked vector of generalized coordinates over the $n$th time window are then defined as
\begin{equation*}
\genstatecollocMatSlabArg{n} \defeq 
\unroll (\genstateDiscreteArg{n,1},\ldots,\genstateDiscreteArg{n,\nstepsArg{n}} ).
\end{equation*}
We now define the weighted space--time residual over the entire window as
\begin{equation*}
\residLMSSlabArg{n} : (\genstatecollocMatySlabArg{};\genstateDiscreteArg{n,0}) \mapsto \begin{bmatrix}
 \sqrt{\quadWeightsLMSScalarArg{n}{1} } \stweightingMatOneArg{n} \residLMSArg{n,1}( \basisspaceArg{n} \genstateyDiscreteArgnt{1} + \stateInterceptArg{n},  \basisspaceArg{n} \genstateDiscreteArg{n,0} + \stateInterceptArg{n},) \\
\vdots \\
 \sqrt{\quadWeightsLMSScalarArg{n}{\nstepsArg{n}} } \stweightingMatOneArg{n} \residLMSArg{n,\nstepsArg{n}}( \basisspaceArg{n} \genstateyDiscreteArgnt{\nstepsArg{n}} + \stateInterceptArg{n}, \ldots,\basisspaceArg{n} \genstateyDiscreteArgnt{\nstepsArg{n} - k^n(\nstepsArg{n}) } + \stateInterceptArg{n}) \\
\end{bmatrix},
\end{equation*}
where $\genstatecollocMatySlabArg{} \defeq \unroll (\genstateyDiscreteArg{1},\ldots,\genstateyDiscreteArg{\nstepsArg{n}} ).$
It is worth noting that, by design,
\begin{equation*}
\objectiveArgLMS{n} \bigg( \basisspaceArg{n} \genstateDiscreteArg{n,0} + \stateInterceptArg{n},\ldots,\basisspaceArg{n} \genstateDiscreteArg{n,\nstepsArg{n}} + \stateInterceptArg{n} \ \bigg) 
\equiv
\bigg[\residLMSSlabArg{n}  (\genstatecollocMatSlabArg{n};\genstateDiscreteArg{n,0}) \bigg]^T \bigg[ \residLMSSlabArg{n}(\genstatecollocMatSlabArg{n};\genstateDiscreteArg{n,0}) \bigg].
\end{equation*} 
Using these definitions, Algorithm~\ref{alg:colloc_gn} presents the standard Gauss--Newton method. The algorithm consists of three fundamental steps: (1) compute the FOM O$\Delta$E residual given the current guess, (2) compute the time window residual-Jacobian and form the normal equations, and (3) solve the normal equations and update the state.

The practical implementation of the Gauss--Newton algorithm requires an efficient method for computing the time window residual-Jacobians. In particular, it is important to note that for linear multistep methods (and many other time marching schemes) the time window residual Jacobian will (almost) be a block-diagonal 
sparse matrix with the sparsity pattern
\begin{equation*}
\frac{\partial \residLMSSlabArg{n}}{\partial \genstatecollocMatySlabArg{}}(\genstatecollocMatSlabArg{n})  =
 \begin{bmatrix*}[l]
\matshapeb & \\%[-5pt]
 \matshapea & \matshapea & \\%[-5pt]
 & \matshapea  & \matshapea & \\%[-5pt]
&  & \ddots & \\%[-5pt]
 & &  & \matshapea &  \matshapea 
\end{bmatrix*}.
% \begin{bmatrix*}[c]
%\matshapeb & \\[-5pt]
% & \hspace{-12pt}\matshapea & \\[-5pt]
% &  & \hspace{-12pt}\matshapea & \\[-5pt]
%&  & \hspace{5pt} \ddots & \\[-5pt]
% & &  & & \hspace{-15 pt} \matshapea 
%\end{bmatrix*}.
\end{equation*}
This sparsity pattern can be leveraged to assemble the time window residual-Jacobian from local residual Jacobians. 
Further, the sparsity pattern can be 
leveraged to efficiently compute the Jacobian matrix-matrix product in the normal equations.
It is also worth noting that the 
normal equations will also (almost) be block-diagonal; this is a fact that can be leveraged to speed up the linear solves at each Gauss--Newton iteration.

@@

The FBSM is an iterative approach that can be used to
solve the two-point boundary value problem. The general process of the FBSM is
as follows: First, the system~\eqref{eq:lspg_continuous} is (numerically) solved forward in
time using an initial guess to the costate to obtain an initial guess to
the generalized coordinates. Next, the adjoint equation~\eqref{eq:lspg_adjoint} is solved \textit{backwards} in time given the
approximation to the generalized coordinates. This gives a new estimate to the
costate, which is then used to again solve the forward problem to obtain
a new estimate of the generalized coordinates. This process is continued until
convergence. Algorithm~\ref{alg:st_iter} outlines the method. The algorithm
contains three parameters: the damping factor $\rho$, the growth factor
$\fbsmGrowth$, and the decay factor $\fbsmDecay$. The damping factor controls the rate at which the costate seen by~\eqref{eq:lspg_continuous} is updated. The
closer $\rho$ is to unity, the faster the resulting algorithm will converge.
For large window sizes, however, too high a value of $\rho$ can lead to an unstable iterative process. 
A proper value of $\rho$ can be
obtained with a line search. The line search presented in Algorithm~\ref{alg:st_iter} adapts the damping factor
according to the objective. Convergence properties of the FBSM method are
presented in Ref.~\cite{McAsey2012ConvergenceOT}. It is shown that, for a small 
enough value of $\rho$, the algorithm will converge.

\begin{algorithm} \caption{\spatialAcronym\ trial subspace: algorithm for the FBSM over the $n$th window.} \label{alg:st_iter} 
\SetKwInOut{Input}{Input}\SetKwInOut{Output}{Output}
\Input{tolerance, $\epsilon$; damping factor, $\rho$; growth factor, $\fbsmGrowth$; decay
factor, $\fbsmDecay$} 
\Output{Stationary point, $\genstateArgnt{n}$ }
\textbf{Online Steps:}\\ 
$\genstate^n_0 \leftarrow \bz$ \Comment{Set initial guess for state} \\
$\controllerArgnt{n} \leftarrow \boldsymbol 0$ \Comment{Set initial
guess for costate}\\ 
$\text{Compute } \genstateArgnt{n}_1 \text{ satisfying } \basisspaceTArg{n} \stweightingMatArg{n}
\basisspaceArg{n} \genstateDotArgnt{n}_1(t)  -  \basisspaceTArg{n} \stweightingMatArg{n}
\velocity(\veloargsromArg{1}) =  \controllerArg{n}{t}$ 
% Eq.~\ref{eq:lspg_continuous} with $\adjoint^n(t) = \controllerArg{n}{t}$ to
% obtain $\genstate_0^{n}(t)$ 
\Comment{Solve~\eqref{eq:lspg_continuous}}\\ 
%Set: $\genstate^n_1 = \genstate$ \Comment{Set state} \\
%$\objectiveArg{n}({\genstate_1^n(t)})$
%\Comment{Evaluate objective function}\\
$i \leftarrow 1$ \Comment{Set counter}\\
\While{$\epsilon \le \int_{\timeStartArg{n}}^{\timeEndArg{n}} \norm{\genstate^n_{i}(t) - \genstate^n_{i-1}(t) }dt $}{
\small{
\begin{multline*}
\text{ Compute } \adjointArgnt{n} \text{ satisfying }
\adjointDotArg{n}{t}  + \basisspaceTArg{n} \bigg[\frac{\partial \velocity}{\partial \stateyDiscrete}(\basisspaceArg{n} \genstate^n_i(t) + \stateInterceptArg{n},t) \bigg]^T \stweightingMatArg{n} \basisspaceArg{n} [\massArg{n}]^{-1} \adjointArg{n}{t}= \\ -\bigg[\basisspaceTArg{n} \bigg[ \frac{\partial \velocity}{\partial \stateyDiscrete} ( \basisspaceArg{n} \genstate^n_i(t) + \stateInterceptArg{n},t) \bigg]^T \stweightingMatArg{n} \bigg( \mathbf{I} -   \basisspaceArg{n} [\massArg{n}]^{-1} \basisspaceTArg{n} \stweightingMatArg{n} \bigg)  \bigg( \basisspaceArg{n} \dot{\genstate}_i^n(t)   -   \velocity( \basisspaceArg{n} \genstate_i^n(t)  +\stateInterceptArg{n},t) \bigg) \bigg] 
\end{multline*} }
\Comment{Solve~\eqref{eq:lspg_adjoint} to obtain guess to costate} \\
%Solve Eq.~\ref{eq:lspg_adjoint} with $\genstate^n(t) = \genstate^n_i(t)$ to obtain $\adjoint^n(t)$ \\
$\controllerArgnt{n}  \leftarrow \rho \controllerArgnt{n} + (1 - \rho) \adjointArgnt{n}$ \Comment{Weighted update to costate}\\
$i \leftarrow i+1$ \Comment{Update counter}\\
$\text{Compute }\genstateArgnt{n}_i \text{ satisfying } \basisspaceTArg{n} \stweightingMatArg{n} \basisspaceArg{n} \genstateDotArgnt{n}_i(t)   -  \basisspaceTArg{n} \stweightingMatArg{n} \velocity(\basisspaceArg{n} \genstate^n_i(t) + \stateInterceptArg{n},t) =  \controllerArg{n}{t} $
\Comment{Solve~\eqref{eq:lspg_continuous}}\\
%$\objectiveArg{n}({\genstate_i^n(t)})$ \Comment{Evaluate objective function}\\
\uIf{ $\objectiveArg{n}({\basisspaceArg{n}\genstate_i^n + \stateInterceptArg{n}\otimes \onesFunctionArg{n}}) \le \objectiveArg{n}({\basisspaceArg{n}\genstate_{i-1}^n + \stateInterceptArg{n}\otimes \onesFunctionArg{n}})$}
{
$\rho \leftarrow \text{min}(\rho \fbsmGrowth,1)$ \Comment{Grow the damping factor}\\
}
\Else{
$\rho \leftarrow \frac{\rho }{ \fbsmDecay}$ \Comment{Shrink the damping factor}\\ 
$\genstate_i^{n} \leftarrow  \genstate_{i-1}^{n}$ \Comment{Reset state to value at previous iteration}
}
}
Return converged solution, $\genstateArgnt{n}= \genstateArgnt{n}_i$
\end{algorithm}
\subsubsection{Numerical considerations for the forward and backward problems}
The FBSM requires solving~\eqref{eq:lspg_continuous} and~\eqref{eq:lspg_adjoint}, both of which are defined at the time-continuous level. 
The numerical implementation of the FBSM requires two main ingredients: (1) temporal discretization schemes for the forward and backward problems and (2) 
an efficient method for computing action of the Jacobian transpose on a vector.

>>>>

\subsection{\spatialAcronym: direct solution approach}\label{sec:direct} 
\KTC{HERE}
\KTC{Does the time grid definted in this section have any relationship to the
time grid used for time discretization? Haven't gotten past HERE below, but
wanted to check (maybe it's made clear later).}
\EP{This time grid becomes the time discretization over each window. At this point in the paper, 
there is no notion of a time grid for \methodAcronym}

Direct approaches solve the optimization problem~\eqref{eq:tclsrm} by
``transcribing" the infinite dimensional optimization problem into a finite
dimensional one by discretizing the state and objective functional in time.
The minimization problem is then reformulated as a (non)linear programming, or
optimization, problem. A variety of direct solution approaches exist, 
including collocation approaches, spectral  methods, and genetic algorithms.

@@

Linear multistep schemes approximate the solution at time instance $\timeWindowArg{n}{i}$ using the previous $k$ time instances.  

@@

\methodAcronym\ sequentially computes the approximate solutions $\approxstateDiscreteArg{n,0},\ldots,\approxstateDiscreteArg{n,\nstepsArg{n}}$, $n=1,\ldots,\nslabs$, as the solution to the minimization problem

@@

\begin{remark}
For the limiting case where $\nstepsArg{n} = 1$ such that the window size is equivalent to the time step, $\DeltaSlabArg{n} \equiv \timeWindowArg{n}{1} - \timeWindowArg{n}{0}$, and uniform 
quadrature weights are used, \methodAcronym\ with \spatialAcronym\ trial subspaces solved via the direct approach recovers LSPG.
\end{remark} 
\subsubsection{Solution to the least-squares problem through the Gauss--Newton method}
Discretization through linear multistep methods (as well as other techniques) 
results in a (non)linear least-squares problem.
A variety of algorithms exist for solving least-squares problems, including trust region approaches, the Gauss–Newton method, and the Levenberg–Marquardt method.  
The numerical experiments presented in this work consider the Gauss--Newton method, and as such we outline this approach here.

To describe the Gauss--Newton method, we first define a ``vectorization" function that takes a tuple of time-local solutions at each time instance and outputs 
a concatenated vector, 
\begin{align*}
 \unroll &\vcentcolon (\stateyDiscreteArg{1},\ldots,\stateyDiscreteArg{m} ) \mapsto \begin{bmatrix} [\stateyDiscreteArg{1}]^T & \ldots & [\stateyDiscreteArg{m}]^T \end{bmatrix}^T  \\
&\vcentcolon \RR{p} \otimes \RR{m} \rightarrow \RR{pm},
\end{align*}
for arbitrary $p$ and $m$. The stacked vector of generalized coordinates over the $n$th time window are then defined as
\begin{equation*}
\genstatecollocMatSlabArg{n} \defeq 
\unroll (\genstateDiscreteArg{n,1},\ldots,\genstateDiscreteArg{n,\nstepsArg{n}} ).
\end{equation*}
We now define the weighted space--time residual over the entire window as
\begin{equation*}
\residLMSSlabArg{n} : (\genstatecollocMatySlabArg{};\genstateDiscreteArg{n,0}) \mapsto \begin{bmatrix}
 \sqrt{\quadWeightsLMSScalarArg{n}{1} } \stweightingMatOneArg{n} \residLMSArg{n,1}( \basisspaceArg{n} \genstateyDiscreteArgnt{1} + \stateInterceptArg{n},  \basisspaceArg{n} \genstateDiscreteArg{n,0} + \stateInterceptArg{n},) \\
\vdots \\
 \sqrt{\quadWeightsLMSScalarArg{n}{\nstepsArg{n}} } \stweightingMatOneArg{n} \residLMSArg{n,\nstepsArg{n}}( \basisspaceArg{n} \genstateyDiscreteArgnt{\nstepsArg{n}} + \stateInterceptArg{n}, \ldots,\basisspaceArg{n} \genstateyDiscreteArgnt{\nstepsArg{n} - k^n(\nstepsArg{n}) } + \stateInterceptArg{n}) \\
\end{bmatrix},
\end{equation*}
where $\genstatecollocMatySlabArg{} \defeq \unroll (\genstateyDiscreteArg{1},\ldots,\genstateyDiscreteArg{\nstepsArg{n}} ).$
It is worth noting that, by design,
\begin{equation*}
\objectiveArgLMS{n} \bigg( \basisspaceArg{n} \genstateDiscreteArg{n,0} + \stateInterceptArg{n},\ldots,\basisspaceArg{n} \genstateDiscreteArg{n,\nstepsArg{n}} + \stateInterceptArg{n} \ \bigg) 
=
\bigg[\residLMSSlabArg{n}  (\genstatecollocMatSlabArg{n};\genstateDiscreteArg{n,0}) \bigg]^T \bigg[ \residLMSSlabArg{n}(\genstatecollocMatSlabArg{n};\genstateDiscreteArg{n,0}) \bigg].
\end{equation*} 
Using these definitions, Algorithm~\ref{alg:colloc_gn} presents the standard Gauss--Newton method. The algorithm consists of three fundamental steps: (1) compute the FOM O$\Delta$E residual given the current guess, (2) compute the Jacobian of the residual over the time window and form the normal equations, and (3) solve the normal equations and update the state.

The practical implementation of the Gauss--Newton algorithm requires an efficient method for computing the Jacobian of the residual over the time window. In particular, it is important to note that for linear multistep methods (and many other time marching schemes) the Jacobian of the residual over the time window will an almost block-diagonal 
sparse matrix with the sparsity pattern
\begin{equation*}
\frac{\partial \residLMSSlabArg{n}}{\partial \genstatecollocMatySlabArg{}}(\genstatecollocMatSlabArg{n})  =
 \begin{bmatrix*}[l]
\matshapeb & \\%[-5pt]
 \matshapea & \matshapea & \\%[-5pt]
 & \matshapea  & \matshapea & \\%[-5pt]
&  & \ddots & \\%[-5pt]
 & &  & \matshapea &  \matshapea 
\end{bmatrix*}.
% \begin{bmatrix*}[c]
%\matshapeb & \\[-5pt]
% & \hspace{-12pt}\matshapea & \\[-5pt]
% &  & \hspace{-12pt}\matshapea & \\[-5pt]
%&  & \hspace{5pt} \ddots & \\[-5pt]
% & &  & & \hspace{-15 pt} \matshapea 
%\end{bmatrix*}.
\end{equation*}
This sparsity pattern can be leveraged to assemble the Jacobian of the residual over the time window from Jacobians of the residual at single time instances. 
Further, the sparsity pattern can be 
leveraged to efficiently compute the Jacobian matrix-matrix product in the normal equations.
It is also worth noting that the 
normal equations will also (almost) be block-diagonal; this is a fact that can be leveraged to speed up the linear solves at each Gauss--Newton iteration.

@@

The FBSM is an iterative approach that can be used to
solve the two-point boundary value problem. The general process of the FBSM is
as follows: First, the system~\eqref{eq:lspg_continuous} is (numerically) solved forward in
time using an initial guess to the costate to obtain an initial guess to
the generalized coordinates. Next, the adjoint equation~\eqref{eq:lspg_adjoint} is solved \textit{backwards} in time given the
approximation to the generalized coordinates. This gives a new estimate to the
costate, which is then used to again solve the forward problem to obtain
a new estimate of the generalized coordinates. This process is continued until
convergence. Algorithm~\ref{alg:st_iter} outlines the method. The algorithm
contains three parameters: the damping factor $\rho \le 1$, the growth factor
$\fbsmGrowth \ge 1$, and the decay factor $\fbsmDecay \ge 1$. The damping factor controls the rate at which the costate seen by~\eqref{eq:lspg_continuous} is updated. The
closer $\rho$ is to unity, the faster the resulting algorithm will converge.
For large window sizes, however, too high a value of $\rho$ can lead to an unstable iterative process. 
A proper value of $\rho$ can be
obtained with a line search. The line search presented in Algorithm~\ref{alg:st_iter} adapts the damping factor
according to the objective. Convergence properties of the FBSM method are
presented in Ref.~\cite{McAsey2012ConvergenceOT}. It is shown that, for a small 
enough value of $\rho$, the algorithm will converge.

\begin{algorithm} \caption{\spatialAcronym\ trial subspace: algorithm for the FBSM over the $n$th window.} \label{alg:st_iter} 
\SetKwInOut{Input}{Input}\SetKwInOut{Output}{Output}
\Input{tolerance, $\epsilon$; damping factor, $\rho \le 1$; growth factor, $\fbsmGrowth \ge 1$; decay
factor, $\fbsmDecay \ge 1$} 
\Output{Stationary point, $\genstateArgnt{n}$ }
\textbf{Online Steps:}\\ 
$\genstate^n_0 \leftarrow \bz$ \Comment{Set initial guess for state} \\
$\controllerArgnt{n} \leftarrow \boldsymbol 0$ \Comment{Set initial
guess for costate}\\ 
$\text{Compute } \genstateArgnt{n}_1 \text{ satisfying } \basisspaceTArg{n} \stweightingMatArg{n}
\basisspaceArg{n} \genstateDotArgnt{n}_1(t)  -  \basisspaceTArg{n} \stweightingMatArg{n}
\velocity(\veloargsromArg{1}) =  \controllerArg{n}{t}$ 
% Eq.~\ref{eq:lspg_continuous} with $\adjoint^n(t) = \controllerArg{n}{t}$ to
% obtain $\genstate_0^{n}(t)$ 
\Comment{Solve~\eqref{eq:lspg_continuous}}\\ 
%Set: $\genstate^n_1 = \genstate$ \Comment{Set state} \\
%$\objectiveArg{n}({\genstate_1^n(t)})$
%\Comment{Evaluate objective function}\\
$i \leftarrow 1$ \Comment{Set counter}\\
\While{$\epsilon \le \int_{\timeStartArg{n}}^{\timeEndArg{n}} \norm{\genstate^n_{i}(t) - \genstate^n_{i-1}(t) }dt $}{
\small{
\begin{multline*}
\text{ Compute } \adjointArgnt{n} \text{ satisfying }
\adjointDotArg{n}{t}  + \basisspaceTArg{n} \bigg[\frac{\partial \velocity}{\partial \stateyDiscrete}(\basisspaceArg{n} \genstate^n_i(t) + \stateInterceptArg{n},t) \bigg]^T \stweightingMatArg{n} \basisspaceArg{n} [\massArg{n}]^{-1} \adjointArg{n}{t}= \\ -\bigg[\basisspaceTArg{n} \bigg[ \frac{\partial \velocity}{\partial \stateyDiscrete} ( \basisspaceArg{n} \genstate^n_i(t) + \stateInterceptArg{n},t) \bigg]^T \stweightingMatArg{n} \bigg( \mathbf{I} -   \basisspaceArg{n} [\massArg{n}]^{-1} \basisspaceTArg{n} \stweightingMatArg{n} \bigg)  \bigg( \basisspaceArg{n} \dot{\genstate}_i^n(t)   -   \velocity( \basisspaceArg{n} \genstate_i^n(t)  +\stateInterceptArg{n},t) \bigg) \bigg] 
\end{multline*} }
\Comment{Solve~\eqref{eq:lspg_adjoint} to obtain guess to costate} \\
%Solve Eq.~\ref{eq:lspg_adjoint} with $\genstate^n(t) = \genstate^n_i(t)$ to obtain $\adjoint^n(t)$ \\
$\controllerArgnt{n}  \leftarrow \rho \controllerArgnt{n} + (1 - \rho) \adjointArgnt{n}$ \Comment{Weighted update to costate}\\
$i \leftarrow i+1$ \Comment{Update counter}\\
$\text{Compute }\genstateArgnt{n}_i \text{ satisfying } \basisspaceTArg{n} \stweightingMatArg{n} \basisspaceArg{n} \genstateDotArgnt{n}_i(t)   -  \basisspaceTArg{n} \stweightingMatArg{n} \velocity(\basisspaceArg{n} \genstate^n_i(t) + \stateInterceptArg{n},t) =  \controllerArg{n}{t} $
\Comment{Solve~\eqref{eq:lspg_continuous}}\\
%$\objectiveArg{n}({\genstate_i^n(t)})$ \Comment{Evaluate objective function}\\
\uIf{ $\objectiveArg{n}({\basisspaceArg{n}\genstate_i^n + \stateInterceptArg{n}\otimes \onesFunctionArg{n}}) \le \objectiveArg{n}({\basisspaceArg{n}\genstate_{i-1}^n + \stateInterceptArg{n}\otimes \onesFunctionArg{n}})$}
{
$\rho \leftarrow \text{min}(\rho \fbsmGrowth,1)$ \Comment{Grow the damping factor}\\
}
\Else{
$\rho \leftarrow \frac{\rho }{ \fbsmDecay}$ \Comment{Shrink the damping factor}\\ 
$\genstate_i^{n} \leftarrow  \genstate_{i-1}^{n}$ \Comment{Reset state to value at previous iteration}
}
}
Return converged solution, $\genstateArgnt{n}= \genstateArgnt{n}_i$
\end{algorithm}
\subsubsection{Numerical considerations for the forward and backward problems}
The FBSM requires solving the systems~\eqref{eq:lspg_continuous} and~\eqref{eq:lspg_adjoint}, both of which are defined at the time-continuous level. 
The numerical implementation of the FBSM requires two main ingredients: (1) temporal discretization schemes for the forward and backward problems and (2) 
an efficient method for computing the action of the Jacobian transpose on a vector.
